---
title: "hw4"
output: html_document
author: Nan Chen
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Access the Yarn cluster
options(scipen=200)
if (!require("pacman"))  
  install.packages("pacman", repos = "http://cran.us.r-project.org/")
library(pacman)
p_load("sparklyr", "dplyr", "ggplot2", "ggmap", "maps", "mapdata", 
           "devtools", "ggrepel", "lubridate", "RColorBrewer")
Sys.setenv(SPARK_HOME="/usr/lib/spark")
config <- spark_config()
sc <- spark_connect(master = "yarn-client", config = config)
```

```{r}
# Connect the Hive tables
flights_tbl <- tbl(sc, 'flights')
airlines_tbl <- tbl(sc, 'airlines')
airports_tbl <- tbl(sc, 'airports')
```

```{r}
# Extract data in 2008
flights2008 <- flights_tbl %>% filter(origin != "Origin") %>%
  select(year, origin, dest) %>% filter(year == 2008)

airports <- airports_tbl %>% select(name, faa, lat, lon)

# Destination with # of flights
dest <- dplyr::left_join(flights2008, airports, by = c("dest" = "faa")) %>%
  group_by(dest) %>% summarise(n = n()) %>% collect() %>% arrange(desc(n)) %>%
  select("airport" = dest, n)

# Origin with # of flights
origin <- dplyr::left_join(flights2008, airports, by = c("origin" = "faa")) %>%
  group_by(origin) %>% summarise(n = n()) %>% collect() %>% arrange(desc(n)) %>%
  select("airport" = origin, n)
```

1. Map the top 10 busiest airports. Size of dots should reflect the number of flights through that destination.  

```{r}
# Add dest and origin together and print out top 10
combine <- bind_rows(dest, origin) %>%
  group_by(airport) %>% summarise(sum = sum(n)) %>% arrange(desc(sum)) %>%
  head(10)

# Pick out airports info for top 10
infotop10 <- airports %>% filter(faa %in% combine$airport) %>% collect()

# Complete info for top 10 airports
top10 <- left_join(combine, infotop10, by = c("airport" = "faa")) %>%
      transmute(airport = airport, totflights = sum, lat = as.numeric(lat), 
                lon = as.numeric(lon), name = name)

usa <- map_data("usa")
states <- map_data("state")
ggplot() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_polygon(data = states, aes(x = states$lon, y = states$lat, fill = region, 
                                  group = group), color = "white") +
  geom_point(data = top10, aes(x = lon, y = lat, size = totflights)) +
  labs(x = "Longitude", y = "Latitude", 
        title = "Top 10 Busiest Airports in 2008", size = "Total Flights") +
  geom_label_repel(data = top10, aes(x = lon, y = lat, label = airport)) +
  coord_fixed(1.3) + guides(fill = FALSE)
```

2. Map the top 10 busiest direct routes. Size of lines should reflect the number of flights through that route.

```{r}
# Top 10 direct routes
route <- flights2008 %>% group_by(origin, dest) %>%
      summarise(n = n()) %>% collect() %>% arrange(desc(n)) %>%
      head(10)
# Destination and origin information
destinfo <- airports %>% filter(faa %in% route$dest) %>% collect() 
dest10 <- left_join(route, destinfo, by = c("dest" = "faa"))

oriinfo <- airports %>% filter(faa %in% route$origin) %>% collect() 
dest_ori <- left_join(dest10, oriinfo, by = c("origin" = "faa"))
dest_ori$scale <- (dest_ori$n / max(dest_ori$n))^10
dest_ori <- dest_ori %>% transmute(dest, lat.to = as.numeric(lat.x),
                      lat.from = as.numeric(lat.y), lon.to = as.numeric(lon.x),
                      lon.from = as.numeric(lon.y), n, scale)
# Airport info
comroute <- c(route$origin, route$dest)
route_airport <- as.data.frame(unique(comroute))
colnames(route_airport) <- "airport"

route_airport <- left_join(route_airport, oriinfo, by = c("airport" = "faa"))
route_airport <- route_airport %>% transmute(lon = as.numeric(lon), 
                          lat = as.numeric(lat), airport, name)

usa <- map_data("usa")
states <- map_data("state")
ggplot() + 
      geom_polygon(data = states, aes(x = long, y = lat, fill = region, 
                            group = group), color = "white") + 
      coord_fixed(1.3) + guides(fill=FALSE) +
      theme(plot.title = element_text(hjust = 0.5)) +
      geom_point(data = route_airport, aes(x = lon, y = lat), colour = "red",
                  alpha = 0.5) + 
      labs(x = "Longitude", y = "Latitude", 
             title = "Top 10 Busiest Routes in 2008") + 
      geom_label_repel(data =  route_airport, aes(x = lon, y = lat), 
                       label =  route_airport$airport) +
      geom_curve(data = dest_ori, aes(x = lon.from, y = lat.from, 
                                      xend = lon.to, yend = lat.to),
                 arrow = arrow(angle = 15, ends = "first", 
                               length = unit(0.2, "cm"), type = "closed"), 
                 size = dest_ori$scale, alpha = 0.7, 
                 curvature = 0.15, inherit.aes = TRUE) 
```

3. LAX:
(a). Reproduce above plot. Visualize and explain some prominent features you observe. For example, what happened at points 1-5?

```{r}
flights_Q3 <- flights_tbl %>% select(year, month, dayofmonth, dayofweek,
      origin, dest) %>% filter(dest == "LAX" | origin == "LAX") %>% 
      filter(year >= 1998 && year <= 2008)
air_traffic <- flights_Q3 %>%
      group_by(year, month, dayofmonth) %>%
      summarise(n = n()) %>%
      collect() %>% 
      mutate(date = make_date(year = year, month = month, day = dayofmonth))

# Create 5 labels for the turning points
dates <- c("2001-09-11", "2004-11-25", "2004-07-04", "2008-01-01", 
           "2001-01-01")
nf <- c(1100, 1000, 1050, 1200, 1200)
num_lab <- c("1", "2", "3", "4", "5")
labels <- data.frame(num_lab, dates, nf)

ggplot(air_traffic, aes(x = date, y = n)) +
  geom_line() +
  labs(x = "Date", y = "Number of Flights", title = "LAX air traffic") +
  geom_label(data = labels, aes(x = as.Date(dates, "%Y-%m-%d"), y = nf, 
                                label = num_lab), size = 5) + coord_fixed(1.4)
```

* `Lable 1` drops dramatically mainly because of the terrorists attack thing. So people were scared of taking flights. `Lable 2` pops up a little bit maybe due to the national day. `Lable 3` pops up maybe because of the holiday. `Lable 4` drops dramatically probably because of the global financial crisis. So people didn't have too much money to spend on travelling around. `Lable 5` pops up a lot maybe due to a good economic environment. So people want to spend more money on travelling.

(b). Visualize and explain seasonal effects.

```{r}
flight_season <- flights_Q3 %>%
      mutate(season = if_else(month %in% c(12, 1, 2), "winter", 
                              if_else(month %in% c(3, 4, 5), "spring", 
                                      if_else(month %in% c(6, 7, 8), "summer",
                                              if_else(month %in% c(9, 10, 11), 
                                                      "fall", "")
                                              )
                                      )
                              )
             ) %>%
      group_by(year, season) %>%
      summarise(n = n()) %>%
      collect()
ggplot(flight_season, aes(x = as.factor(year), y = n, fill = season)) + 
  geom_col(stat = "identity", position = "dodge", width = 0.7) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x = "Year", y = "Number of Flights",
       title = "LAX Air Traffic by Season", fill = "Season") + 
  scale_fill_brewer(palette = "Set2")
```

(c). Visualize and explain weekly effects.

```{r}
flight_week <- flights_Q3 %>%
  group_by(year, dayofweek) %>%
  summarise(n = n()) %>%
  collect()
ggplot(flight_week, aes(x = as.factor(year), y = n, 
                        fill = factor(dayofweek, labels = c("Sunday", "Monday", "Tuesday", "Wednesday",  "Thursday", "Friday", "Saturday")))) + 
  geom_col() + 
  labs(x = "Year", y = "Number of Flights", fill = "Day",
       title = "LAX Air Traffic by Day of the Week") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_brewer(palette = "Set2")
```

(d). Map top 10 destinations from LAX. Size of dots should reflect the number of flights from LAX to that destination.

```{r}
LAX <- flights2008 %>% filter(origin == "LAX") %>% group_by(dest) %>%
      summarise(n = n()) %>% arrange(desc(n)) %>% collect()
destLAX <- airports %>% filter(faa %in% LAX$dest) %>% collect()
oriLAX <- left_join(destLAX, LAX, by = c("faa" = "dest")) %>%
  mutate(lat = as.numeric(lat), lon = as.numeric(lon)) %>% 
  arrange(desc(n)) %>% collect() %>%
  head(10)

ggplot() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_polygon(data = states, aes(x = states$lon, y = states$lat, fill = region, 
                                  group = group), color = "white") +
  geom_point(data = oriLAX, aes(x = lon, y = lat, size = n)) +
  labs(x = "Longitude", y = "Latitude", 
        title = "Top 10 Destinations from LAX", size = "Total Flights") +
  geom_label_repel(data = oriLAX, aes(x = lon, y = lat, label = faa)) +
  coord_fixed(1.3) + guides(fill = FALSE)
```

4. Build a predictive model for the arrival delay (`arrdelay`) of flights flying from LAX. Use the same filtering criteria as in the [lecture notes](http://hua-zhou.github.io/teaching/biostatm280-2018winter/slides/12-sparklyr/sparklyr-flights.html) to construct training and validation sets. You are allowed to use a maximum of 5 predictors. The prediction performance of your model on the validation data set will be an important factor for grading this question.

```{r}
model_data <- flights_tbl %>% filter(origin == "LAX") %>%
    filter(!is.na(arrdelay) & !is.na(depdelay) & !is.na(distance)) %>%
    filter(depdelay > 15 & depdelay < 240) %>%
    filter(arrdelay > -60 & arrdelay < 360) %>%
    filter(year >= 2003 & year <= 2007) %>%
    left_join(airlines_tbl, by = c("uniquecarrier" = "code")) %>%
    select(year, month, arrdelay, depdelay, distance, uniquecarrier, 
           description)

# Summarize data by carrier
model_data %>%
  group_by(uniquecarrier) %>%
  summarize(description = min(description, na.rm = TRUE), 
            arrdelay = mean(arrdelay, na.rm = TRUE), 
            distance = mean(distance, na.rm = TRUE), 
            depdelay = mean(depdelay, na.rm = TRUE)) %>%
  select(description, arrdelay, distance, depdelay) %>%
  arrange(arrdelay)

# Partition the data into training and validation sets
model_partition <- model_data %>% 
  sdf_partition(train = 0.8, valid = 0.2, seed = 5555)

# Fit a linear model
ml1 <- model_partition$train %>%
    ml_linear_regression(arrdelay ~ distance + depdelay + uniquecarrier)

# Summarize the linear model
summary(ml1)

# Get mse for valid
sdf_predict(ml1, model_partition$valid) %>%
  mutate(res = prediction - arrdelay) %>%
  summarise(mse = mean(res^2, na.rm = TRUE))
```

* Root Mean Squared Error for train data is `13.33`; MSE for valid data is `179.91`.

```{r}
# Assess model performance
# Calculate average arrdelay by predicted decile
model_deciles <- lapply(model_partition, function(x) {
    sdf_predict(ml1, x) %>%
      mutate(decile = ntile(desc(prediction), 10)) %>%
      group_by(decile) %>%
      summarize(arrdelay = mean(arrdelay, na.rm = TRUE)) %>%
      select(decile, arrdelay) %>%
      collect()
  })

# Create a summary dataset for plotting
deciles <- rbind(
  data.frame(data = 'train', model_deciles$train),
  data.frame(data = 'valid', model_deciles$valid),
  make.row.names = FALSE
)
# Plot average arrdelay by predicted decile
deciles %>%
  ggplot(aes(factor(decile), arrdelay, fill = data)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  labs(title = 'Average arrdelay by predicted decile', x = 'Decile', 
       y = 'Minutes')

# Predict performance
# Select data from an out of time sample
data_2008 <- flights_tbl %>% filter(origin == "LAX") %>%
  filter(!is.na(arrdelay) & !is.na(depdelay) & !is.na(distance)) %>%
  filter(depdelay > 15 & depdelay < 240) %>%
  filter(arrdelay > -60 & arrdelay < 360) %>%
  filter(year == 2008) %>%
  left_join(airlines_tbl, by = c("uniquecarrier" = "code")) %>%
  select(year, month, arrdelay, depdelay, distance, uniquecarrier, description, 
         origin, dest)

# Summarize data by carrier
carrier <- sdf_predict(ml1, data_2008) %>%
  group_by(description) %>%
  summarize(arrdelay = mean(arrdelay, na.rm = TRUE), 
            prediction = mean(prediction, na.rm = TRUE), freq = n()) %>%
  filter(freq > 100) %>%
  collect()

# Get mse for test
sdf_predict(ml1, data_2008) %>%
  mutate(res = prediction - arrdelay) %>%
  summarise(mse = mean(res^2, na.rm = TRUE))

# Plot actual arrdelay and predicted arrdelay by airline carrier
ggplot(carrier, aes(arrdelay, prediction)) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_point(alpha = 0.75, color = 'red', shape = 3) +
  geom_abline(intercept = 0, slope = 1, alpha = 0.15, color = 'blue') +
  geom_text(aes(label = substr(description, 1, 20)), size = 3, alpha = 0.75, 
            vjust = -1) +
  labs(title='Average Arrival Delay Forecast', x = 'Actual', y = 'Predicted')
```

* MSE for test data is `203.03`.

5. Visualize and explain any other information you want to explore.

```{r}
SEA <- flights2008 %>% filter(origin == "SEA") %>% group_by(dest) %>%
      summarise(n = n()) %>% arrange(desc(n)) %>% collect()
destSEA <- airports %>% filter(faa %in% SEA$dest) %>% collect()
oriSEA <- left_join(destSEA, SEA, by = c("faa" = "dest")) %>%
  mutate(lat = as.numeric(lat), lon = as.numeric(lon)) %>% 
  arrange(desc(n)) %>% collect() %>%
  head(10)

ggplot() +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_polygon(data = states, aes(x = states$lon, y = states$lat, fill = region, 
                                  group = group), color = "white") +
  geom_point(data = oriSEA, aes(x = lon, y = lat, size = n)) +
  labs(x = "Longitude", y = "Latitude", 
        title = "Top 10 Destinations from SEA", size = "Total Flights") +
  geom_label_repel(data = oriSEA, aes(x = lon, y = lat, label = faa)) +
  coord_fixed(1.3) + guides(fill = FALSE)
```







