---
title: "hw3"
author: "Nan Chen"
date: "2/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Q1. LA City Employee Payroll
Shiny App Website: https://nan-chen2018.shinyapps.io/la-payroll2017/

##Q2. LA City Parking War
```{r}
    if (!"DBI" %in% rownames(installed.packages()))  
      install.packages("DBI", repos = "http://cran.rstudio.com/")
    if (!"RSQLite" %in% rownames(installed.packages()))  
      install.packages("RSQLite", repos = "http://cran.rstudio.com/") 
    if (!"tidyverse" %in% rownames(installed.packages()))  
      install.packages("tidyverse", repos = "http://cran.rstudio.com/")
    if (!"dplyr" %in% rownames(installed.packages()))  
      install.packages("dplyr", repos = "http://cran.rstudio.com/")
suppressMessages(library("DBI"))
suppressMessages(library("RSQLite"))
suppressMessages(library("tidyverse"))
suppressMessages(library("dplyr"))

```

```{r}
# read in data from SQLite
con <- dbConnect(RSQLite::SQLite(), 
                "/home/m280-data/la_parking/LA_Parking_Citations_Extra.sqlite")

# show dataset name in con, "latix"
dbListTables(con)

# deposit latix to R
latix <- dplyr::tbl(con, "latix")

```

1. How many tickets are in this data set? Which time period do these tickets span? Which years have most data?
```{r}
head(latix, n = , width = Inf)
latix %>% filter(!is.na(Ticket_number)) %>%
  count()
```
* There are 4044338 tickets in this dataset after removing NA values. 

```{r}
# remove missing dates
date <- latix %>% filter(!is.na(Issue_Year), !is.na(Issue_Month), 
                         !is.na(Issue_Day), !is.na(Issue_Hour),
                         !is.na(Issue_Minute), !is.na(Issue_Wday)) 
# tickets time span
date %>% distinct(Issue_Year, Issue_Month, Issue_Day, Issue_Hour, 
                  Issue_Minute) %>%
  arrange(desc(Issue_Year), desc(Issue_Month), desc(Issue_Day), 
          desc(Issue_Hour), desc(Issue_Minute)) %>%
  collect() %>% slice(c(1, n()))
```
* These pictures span from 2010-04-27 21:40 to 2017-12-30 01:41.

```{r}
date %>% filter(!is.na(Issue_Year)) %>% group_by(Issue_Year) %>%
  summarise(count = n()) %>% arrange(desc(count)) %>% head(1)
```
* Year 2015 has most data. 

2. When (which hour, weekday, month day, and month) are you most likely to get a ticket and when are you least likely to get a ticket?
```{r}
# hour
date %>% group_by(Issue_Hour) %>% 
  summarise(count = n()) %>% arrange(desc(count)) %>%
  collect() %>% slice(c(1, n()))
```
* It most likely to get a ticket at hour '12' and least likely to get a ticket at hour '5'. 

```{r}
# weekday
date %>% group_by(Issue_Wday) %>% 
  summarise(count = n()) %>% arrange(desc(count)) %>%
  collect() %>% slice(c(1, n()))
```
* It most likely to get a ticket on 'Tuesday' and least likely to get a ticket on 'Saturday'. 

```{r}
# day
date %>% group_by(Issue_Day) %>% 
  summarise(count = n()) %>% arrange(desc(count)) %>%
  collect() %>% slice(c(1, n()))
```
* It most likely to get a ticket on day '22' and least likely to get a ticket on day '31'. 

```{r}
# month
date %>% group_by(Issue_Month) %>% 
  summarise(count = n()) %>% arrange(desc(count)) %>%
  collect() %>% slice(c(1, n()))
```
* It most likely to get a ticket in 'March' and least likely to get a ticket in 'November'. 

3. Which car makes received most citations?
```{r}
# remove NA
car <- latix %>% filter(!is.na(Make)) 
car %>% group_by(Make) %>%
  summarise(count = n()) %>% arrange(desc(count)) %>% head(1)
```
* 'TOYOTA' received most citations. 

4. How many different colors of cars were ticketed? Which color attracted most tickets?
```{r}
# remove NA
color <- latix %>% filter(!is.na(Color))
# number of colors
color1 <- color %>% group_by(Color) %>% 
  summarise(count = n()) %>% arrange(desc(count)) 
summarise(color1, n())
# color that most attracted tickets
color1 %>% head(1)
```
* There are '65' different colors of cars were ticketed. The 'Black' one attracted most tickets.

















