---
title: "HW3"
author: "Nan Chen"
date: "2/28/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

####Q1. LA City Employee Payroll
Shiny App Website: https://nan-chen2018.shinyapps.io/la-payroll2017/

####Q2. LA City Parking War
```{r}
if (!"DBI" %in% rownames(installed.packages()))  
      install.packages("DBI", repos = "http://cran.rstudio.com/")
if (!"RSQLite" %in% rownames(installed.packages()))  
      install.packages("RSQLite", repos = "http://cran.rstudio.com/") 
if (!"tidyverse" %in% rownames(installed.packages()))  
      install.packages("tidyverse", repos = "http://cran.rstudio.com/")
if (!"dplyr" %in% rownames(installed.packages()))  
      install.packages("dplyr", repos = "http://cran.rstudio.com/")
library(DBI) 
library(RSQLite)
library(tidyverse)
library(dplyr)
```

```{r}
# read in data from SQLite
con <- dbConnect(RSQLite::SQLite(), 
                "/home/m280-data/la_parking/LA_Parking_Citations_Extra.sqlite")

# show dataset name in con, "latix"
dbListTables(con)

# deposit latix to R
latix <- dplyr::tbl(con, "latix")
```

1. How many tickets are in this data set? Which time period do these tickets span? Which years have most data?

```{r}
latix %>% filter(!is.na(Ticket_number)) %>%
  count()
```

* There are `4044338` tickets in this dataset after removing NA values. 

```{r}
# remove missing dates
date <- latix %>% filter(!is.na(Issue_Year), !is.na(Issue_Month), 
                         !is.na(Issue_Day), !is.na(Issue_Hour),
                         !is.na(Issue_Minute), !is.na(Issue_Wday)) 
# tickets time span
date %>% distinct(Issue_Year, Issue_Month, Issue_Day, Issue_Hour, 
                  Issue_Minute) %>%
  arrange(desc(Issue_Year), desc(Issue_Month), desc(Issue_Day), 
          desc(Issue_Hour), desc(Issue_Minute)) %>%
  collect() %>% slice(c(1, n()))
```

* These pictures span from `2010-04-27 21:40` to `2017-12-30 01:41`.

```{r}
date %>% filter(!is.na(Issue_Year)) %>% group_by(Issue_Year) %>%
  summarise(count = n()) %>% arrange(desc(count)) %>% head(1)
```

* Year `2015` has most data. 

2. When (which hour, weekday, month day, and month) are you most likely to get a ticket and when are you least likely to get a ticket?

```{r}
# hour
date %>% group_by(Issue_Hour) %>% 
  summarise(count = n()) %>% arrange(desc(count)) %>%
  collect() %>% slice(c(1, n()))
```

* It most likely to get a ticket at hour `12` and least likely to get a ticket at hour `5`. 

```{r}
# weekday
date %>% group_by(Issue_Wday) %>% 
  summarise(count = n()) %>% arrange(desc(count)) %>%
  collect() %>% slice(c(1, n()))
```

* It most likely to get a ticket on `Tuesday` and least likely to get a ticket on `Saturday`. 

```{r}
# day
date %>% group_by(Issue_Day) %>% 
  summarise(count = n()) %>% arrange(desc(count)) %>%
  collect() %>% slice(c(1, n()))
```

* It most likely to get a ticket on day `22` and least likely to get a ticket on day `31`. 

```{r}
# month
date %>% group_by(Issue_Month) %>% 
  summarise(count = n()) %>% arrange(desc(count)) %>%
  collect() %>% slice(c(1, n()))
```

* It most likely to get a ticket in `March` and least likely to get a ticket in `November`. 

3. Which car makes received most citations?

```{r}
# remove NA
car <- latix %>% filter(!is.na(Make)) 
car %>% group_by(Make) %>%
  summarise(count = n()) %>% arrange(desc(count)) %>% head(1)
```

* `TOYOTA` received most citations. 

4. How many different colors of cars were ticketed? Which color attracted most tickets?

```{r}
# remove NA
color <- latix %>% filter(!is.na(Color))
# number of colors
color1 <- color %>% group_by(Color) %>% 
  summarise(count = n()) %>% arrange(desc(count)) 
summarise(color1, n())
# color that most attracted tickets
color1 %>% head(1)
```

* There are `65` different colors of cars were ticketed. The `Black` one attracted most tickets.

5. What are the most common ticket types?

```{r}
type <- latix %>% filter(!is.na(Violation_Description))
type %>% group_by(Violation_Description) %>% summarise(count = n()) %>%
  arrange(desc(count)) %>% head(1)
```

* The most common ticket type is `No Park/Street Clean`.

6. How much money was collected on parking tickets in 2015 and 2016?

```{r}
latix %>% select(Fine_amount, Issue_Year) %>%
  filter(!is.na(Fine_amount), !is.na(Issue_Year), Issue_Year >= 2015, 
         Issue_Year <= 2016) %>% 
  summarise(total = sum(Fine_amount, na.rm = TRUE))
```

* The amount of `$274242930` was collected on parking tickets between 2015 and 2016.

7. Which state has the highest fine amount?

```{r}
latix %>% select(RP_State_Plate, Fine_amount) %>% group_by(RP_State_Plate) %>%
  summarise(total = sum(Fine_amount, na.rm = TRUE)) %>%
  arrange(desc(total)) %>% head(1)
```

* `CA` has the highest fine amount.















